---
title: "W10-Dashboards Assignment"
subtitle: "ODI Grips Instagram Dashboard Projects"
author: "Jeffrey Hsu"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: true
    embed-resources: true
    theme: cosmo
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

# Overview

This assignment includes two dashboard projects built from the ODI Grips Instagram dataset. Project 1 is a static Quarto dashboard designed to summarize overall Instagram performance. Project 2 is an interactive Shiny app designed to let users explore how performance changes across date ranges, content dimensions, and marketing metrics. This report includes the project descriptions, data descriptions, source code, findings, chart interpretations, and deployment links for both projects.

# Project 1

## Project Description

Project 1 is a static dashboard built in Quarto. The goal of this dashboard is to present a clear overview of ODI Grips’ Instagram performance to an audience that may not have prior knowledge of the brand or dataset. The story focuses on overall account performance, monthly engagement trends, media format performance, caption strategy performance, and the top-performing posts.

## Data Description

The dataset contains ODI Grips Instagram post-level performance data. Key variables used in this dashboard include:

- `post_date`: date of the Instagram post
- `media_type`: type of media used in the post
- `caption_type`: category of caption strategy used
- `reach`: number of accounts reached
- `likes`, `comments`, `shares`, `saves`: engagement metrics
- `follows`: number of follows generated
- `engagement_rate`: engagement relative to reach or audience

The dashboard uses these variables to summarize performance across time and content strategy.

````qmd
## R Code

````r
---
title: "ODI Grips Instagram Performance Dashboard"
subtitle: "Static Dashboard – Project 1"
format:
  dashboard:
    theme: cosmo
    orientation: rows
    scrolling: true
execute:
  echo: false
  warning: false
  message: false
  freeze: auto
editor: source
---

```{r}
library(tidyverse)
library(lubridate)
library(scales)
library(gt)
library(janitor)

theme_set(theme_minimal())

odi <- read_csv("@odi_grips Instagram Content Data Final Version - V2.csv") |>
  clean_names() |>
  mutate(
    post_date = as.Date(post_date),
    year_month = floor_date(post_date, unit = "month"),
    media_type = as.factor(media_type),
    caption_type = as.factor(caption_type)
  )

total_posts <- nrow(odi)
avg_reach <- mean(odi$reach, na.rm = TRUE)
avg_engagement <- mean(odi$engagement_rate, na.rm = TRUE)
total_follows <- sum(odi$follows, na.rm = TRUE)
```

## Row {height=18%}

### Total Posts

```{r}
#| content: valuebox
#| icon: hash
#| color: primary
total_posts
```

### Average Reach

```{r}
#| content: valuebox
#| icon: eye
#| color: success
comma(round(avg_reach, 0))
```

### Average Engagement Rate

```{r}
#| content: valuebox
#| icon: graph-up
#| color: warning
percent(avg_engagement, accuracy = 0.01)
```

### Total Follows

```{r}
#| content: valuebox
#| icon: people
#| color: info
comma(total_follows)
```

## Row {height=42%}

### Engagement Trend Over Time

```{r}
monthly_perf <- odi |>
  group_by(year_month) |>
  summarise(
    avg_engagement_rate = mean(engagement_rate, na.rm = TRUE),
    avg_reach = mean(reach, na.rm = TRUE),
    posts = n(),
    .groups = "drop"
  )

ggplot(monthly_perf, aes(x = year_month, y = avg_engagement_rate)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  scale_x_date(date_labels = "%b %Y") +
  labs(
    title = "Average Engagement Rate by Month",
    x = "Month",
    y = "Average Engagement Rate"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Reach by Media Type

```{r}
media_summary <- odi |>
  group_by(media_type) |>
  summarise(
    avg_reach = mean(reach, na.rm = TRUE),
    avg_engagement_rate = mean(engagement_rate, na.rm = TRUE),
    posts = n(),
    .groups = "drop"
  ) |>
  arrange(desc(avg_reach))

ggplot(media_summary, aes(x = reorder(media_type, avg_reach), y = avg_reach, fill = media_type)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = comma_format()) +
  labs(
    title = "Average Reach by Media Type",
    x = "Media Type",
    y = "Average Reach"
  )
```

## Row {height=40%}

### Caption Strategy Performance

```{r}
caption_summary <- odi |>
  group_by(caption_type) |>
  summarise(
    avg_engagement_rate = mean(engagement_rate, na.rm = TRUE),
    avg_reach = mean(reach, na.rm = TRUE),
    posts = n(),
    .groups = "drop"
  ) |>
  arrange(desc(avg_engagement_rate))

ggplot(caption_summary, aes(x = reorder(caption_type, avg_engagement_rate), y = avg_engagement_rate, fill = caption_type)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "Average Engagement Rate by Caption Type",
    x = "Caption Type",
    y = "Average Engagement Rate"
  )
```

### Top Performing Posts

```{r}
top_posts <- odi |>
  select(
    post_date,
    media_type,
    caption_type,
    reach,
    likes,
    shares,
    comments,
    saves,
    follows,
    engagement_rate
  ) |>
  arrange(desc(engagement_rate)) |>
  slice_head(n = 10)

top_posts |>
  mutate(
    post_date = format(post_date, "%Y-%m-%d"),
    engagement_rate = percent(engagement_rate, accuracy = 0.01)
  ) |>
  gt() |>
  tab_header(
    title = "Top 10 Posts by Engagement Rate"
  ) |>
  fmt_number(
    columns = c(reach, likes, shares, comments, saves, follows),
    decimals = 0
  )
```

## Row

### Key Takeaways

ODI Grips’ Instagram performance is uneven across content types, with certain media formats and caption strategies consistently outperforming others. Engagement also changes over time, which suggests content timing and content structure both matter.

The dashboard suggests that ODI should prioritize the formats and caption strategies that already show strong reach and engagement rather than spreading effort evenly across all post types.
